---
title: "Exam_question3"
author: "Natalia Giezek"
date: '2022-07-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
### Clear memory
rm(list = ls())

library(tidyverse)
### Load Apollo library
library(apollo)

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_Exam_question3",
  modelDescr      = "Simple MNL model",
  indivID         = "id", #name of id column in our data
  outputDirectory = "output"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

# Make sure that working directory is set to script file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Locate auto.RData in the same folder as this script file
load("shoedata2022.rdata")
database = shoedata
```
a)
```{r}
table_of_choices <-database %>%
  group_by(choice) %>% 
  summarise(n = n())
table_of_choices

table_of_means <- database %>%
  group_by(choice) %>%
summarise(av_p0 = 0, av_p1 = mean(p1), av_p2 = mean(p2), av_p3 = mean(p3), av_p4 = mean(p4), av_p5 = mean(p5))
table_of_means
```
The table "table_of_means" shows the average prices of both chosen and not chosen alternatives, all of which are important; as the consumers not always choose the cheapest alternatives, maybe there are factors other than price which are important to them.

b) A standard multinomial logit model
```{r}
# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, assuming there are no alternative-specific constants in this particular model, as a given utility function does not include them
apollo_beta=c(
 b_1 = 0,
 b_2 = 0,
 b_price =0
)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta
apollo_fixed = c()

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities <-
  function(apollo_beta,
           apollo_inputs,
           functionality = "estimate") {
    ### Attach inputs and detach after function exit
    apollo_attach(apollo_beta, apollo_inputs)
    on.exit(apollo_detach(apollo_beta, apollo_inputs))
    
    ### Create list of probabilities P
    P = list()
    
    ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
    V = list()
    V[["oo"]] =  b_1 * X10 + b_2*X20 - b_price*p0
    V[["opt1"]] =  b_1 * X11 + b_2*X21 - b_price*p1
    V[["opt2"]] =  b_1 * X12 + b_2*X22 - b_price*p2
    V[["opt3"]] =  b_1 * X13 + b_2*X23 - b_price*p3
    V[["opt4"]] =  b_1 * X14 + b_2*X24 - b_price*p4
    V[["opt5"]] =  b_1 * X15 + b_2*X25 - b_price*p5
    
    ### Define settings for MNL model component
    mnl_settings = list(
      alternatives  = c("oo"=0,"opt1" = 1, "opt2" = 2, "opt3" = 3,"opt4" = 4, "opt5" = 5), 
      avail         = list(oo = 1,opt1 = 1, opt2 = 1, opt3 = 1,opt4 = 1, opt5 = 1),
      choiceVar     = choice,
      utilities     = V
    )
    
    ### Compute probabilities using MNL model
    P[["model"]] = apollo_mnl(mnl_settings, functionality)
    
    ### Prepare and return outputs of function
    P = apollo_prepareProb(P, apollo_inputs, functionality)
    return(P)
  }

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model, modelOutput_settings = list(printClassical = FALSE))

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model, saveOutput_settings = list(printClassical = FALSE))
```
As |t| >2, so the estimates are significant. All have positive signs, as expected, as the price entered utility specification with a minus sign, so the positive coefficient is expected, the higher the price the lower the utility.
Therefore, improvements in sole quality, upper durability, and reduction in price have a positive effects on utility.
Willingness to pay:
$$ b_1X_{1j} + b_2X_{2j} - \alpha p_j = b_1X_{1j} + b_2(X_{2j} + 1) - \alpha (p_j + WTP) $$
$$   - \alpha p_j = b_2 - \alpha p_j  - \alpha WTP $$
$$   \alpha WTP = b_2  $$
$$   WTP = \frac{b_2}{\alpha}  $$
Here $$ \alpha = b_{price} $$
```{r}
# Willingness to pay
b_price_hat <- model$estimate[names(model$estimate) == "b_price"]
model$estimate / b_price_hat

deltaMethod_settings = list(
  expression = c(
    WTP_sole = "b_2/b_price"))

apollo_deltaMethod(model, deltaMethod_settings)

```
According to the rule of thumb, when |t| > 2, then it is significant, which is the case here. The WTP for one-unit improvement in the sole quality is 0.5458. However, it is hard to interpret what does this one- unit quality improvement actually mean.

c) For option 3 and option 5, introduce "sustainable" variable
```{r}
### Clear memory
rm(list = ls())

library(tidyverse)
### Load Apollo library
library(apollo)

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_Exam_question3c",
  modelDescr      = "Simple MNL model",
  indivID         = "id", #name of id column in our data
  outputDirectory = "output"
)

# ################################################################# #
#### LOAD DATA AND APPLY ANY TRANSFORMATIONS                     ####
# ################################################################# #

# Make sure that working directory is set to script file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Locate auto.RData in the same folder as this script file
load("shoedata2022.rdata")
database = shoedata
database$sustainable <- ifelse(database$choice == 3 | database$choice == 5, 1,0)
```
```{r}
# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(
 b_1 = 0,
 b_2 = 0,
 b_price =0,
 b_sus = 0
)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c()

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities <-
  function(apollo_beta,
           apollo_inputs,
           functionality = "estimate") {
    ### Attach inputs and detach after function exit
    apollo_attach(apollo_beta, apollo_inputs)
    on.exit(apollo_detach(apollo_beta, apollo_inputs))
    
    ### Create list of probabilities P
    P = list()
    
    ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
    V = list()
    V[["oo"]] =  b_1 * X10 + b_2*X20 - b_price*p0 + b_sus*sustainable
    V[["opt1"]] =  b_1 * X11 + b_2*X21 - b_price*p1 + b_sus*sustainable
    V[["opt2"]] =  b_1 * X12 + b_2*X22 - b_price*p2 + b_sus*sustainable
    V[["opt3"]] =  b_1 * X13 + b_2*X23 - b_price*p3 + b_sus*sustainable
    V[["opt4"]] =  b_1 * X14 + b_2*X24 - b_price*p4 + b_sus*sustainable
    V[["opt5"]] =  b_1 * X15 + b_2*X25 - b_price*p5 + b_sus*sustainable
    
    ### Define settings for MNL model component
    mnl_settings = list(
      alternatives  = c("oo"=0,"opt1" = 1, "opt2" = 2, "opt3" = 3,"opt4" = 4, "opt5" = 5), #quotation marks are important
      avail         = list("oo"= TRUE,"opt1" = TRUE, "opt2" = TRUE, "opt3" = TRUE,"opt4"= TRUE, "opt5" = TRUE),
      choiceVar     = choice,
      utilities     = V
    )
    
    ### Compute probabilities using MNL model
    P[["model"]] = apollo_mnl(mnl_settings, functionality)
    
    ### Prepare and return outputs of function
    P = apollo_prepareProb(P, apollo_inputs, functionality)
    return(P)
  }

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model, modelOutput_settings = list(printClassical = FALSE))

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model, saveOutput_settings = list(printClassical = FALSE))
```
Sustainability seems to have no effect, as it is not significantly different from 0.

d) Nested logit
```{r}
rm(list = ls())
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(tidyverse)
library(apollo)
load("shoedata2022.rdata")
database = shoedata
### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_Nested_Exam_3d",
  modelDescr      = "Simple MNL model with Nests",
  indivID         = "id",
  outputDirectory = "output"
)

# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta = c(
 b_1 = 0,
 b_2 = 0,
 b_price =0,
 lambda_nest1 = 1
)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c() #doesn't matter what constant to fix. Fixing gc to match the lecture slides

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities = function(apollo_beta,
                                apollo_inputs,
                                functionality = "estimate") {
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  
    V[["oo"]] =  b_1 * X10 + b_2*X20 - b_price*p0
    V[["opt1"]] =  b_1 * X11 + b_2*X21 - b_price*p1
    V[["opt2"]] =  b_1 * X12 + b_2*X22 - b_price*p2
    V[["opt3"]] =  b_1 * X13 + b_2*X23 - b_price*p3
    V[["opt4"]] =  b_1 * X14 + b_2*X24 - b_price*p4
    V[["opt5"]] =  b_1 * X15 + b_2*X25 - b_price*p5  

  ### Specify nests for NL model
  nlNests = list(root = 1,
                 nest1 = lambda_nest1
                 )
  
  ### Specify tree structure for NL model
  nlStructure = list()
  nlStructure[["root"]] = c("oo", "nest1")
  nlStructure[["nest1"]] = c("opt1","opt2","opt3","opt4", "opt5")
  
  ### Define settings for NL model
  nl_settings <- list(
    alternatives = c(
      oo =   0,
      opt1 = 1,
      opt2 = 2,
      opt3 = 3,
      opt4 = 4,
      opt5 = 5
    ),
    avail = list(
      oo =   1,
      opt1 = 1,
      opt2 = 1,
      opt3 = 1,
      opt4 = 1,
      opt5 = 1
    ),
    choiceVar = choice,
    utilities = V,
    nlNests      = nlNests,
    nlStructure  = nlStructure
  )
  
  ### Compute probabilities using NL model
  P[["model"]] = apollo_nl(nl_settings, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta,
                        apollo_fixed,
                        apollo_probabilities,
                        apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model, modelOutput_settings = list(printT1=1))
# modelOutput_settings = list(printT1=1) adds t-values for the test whether the coefficient is significantly different from 1 - that's relevant for the lambdas! A lambda equal to 1 would mean that there actually is no nest.

apollo_lrTest(model, "MNL_Exam_question3")

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)
```
e) Modified nests
```{r}
rm(list = ls())
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(tidyverse)
library(apollo)
load("shoedata2022.rdata")
database = shoedata
### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_Nested_Exam_3e",
  modelDescr      = "Simple MNL model with Nests",
  indivID         = "id",
  outputDirectory = "output"
)

# ################################################################# #
#### DEFINE MODEL PARAMETERS                                     ####
# ################################################################# #

### Vector of parameters, including any that are kept fixed in estimation
apollo_beta = c(
 b_1 = 0,
 b_2 = 0,
 b_price =0,
 lambda_nest1 = 1,
 lambda_nest2 = 1
)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c() #doesn't matter what constant to fix. Fixing gc to match the lecture slides

# ################################################################# #
#### GROUP AND VALIDATE INPUTS                                   ####
# ################################################################# #

apollo_inputs = apollo_validateInputs()

# ################################################################# #
#### DEFINE MODEL AND LIKELIHOOD FUNCTION                        ####
# ################################################################# #

apollo_probabilities = function(apollo_beta,
                                apollo_inputs,
                                functionality = "estimate") {
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  
    V[["oo"]] =  b_1 * X10 + b_2*X20 - b_price*p0
    V[["opt1"]] =  b_1 * X11 + b_2*X21 - b_price*p1
    V[["opt2"]] =  b_1 * X12 + b_2*X22 - b_price*p2
    V[["opt3"]] =  b_1 * X13 + b_2*X23 - b_price*p3
    V[["opt4"]] =  b_1 * X14 + b_2*X24 - b_price*p4
    V[["opt5"]] =  b_1 * X15 + b_2*X25 - b_price*p5  

  ### Specify nests for NL model
  nlNests = list(root = 1,
                 nest1 = lambda_nest1,
                 nest2 = lambda_nest2
                 )
  
  ### Specify tree structure for NL model
  nlStructure = list()
  nlStructure[["root"]] = c("oo", "nest1", "nest2")
  nlStructure[["nest1"]] = c("opt1","opt2","opt3")
  nlStructure[["nest2"]] = c("opt4", "opt5")
  ### Define settings for NL model
  nl_settings <- list(
    alternatives = c(
      oo =   0,
      opt1 = 1,
      opt2 = 2,
      opt3 = 3,
      opt4 = 4,
      opt5 = 5
    ),
    avail = list(
      oo =   1,
      opt1 = 1,
      opt2 = 1,
      opt3 = 1,
      opt4 = 1,
      opt5 = 1
    ),
    choiceVar = choice,
    utilities = V,
    nlNests      = nlNests,
    nlStructure  = nlStructure
  )
  
  ### Compute probabilities using NL model
  P[["model"]] = apollo_nl(nl_settings, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta,
                        apollo_fixed,
                        apollo_probabilities,
                        apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model, modelOutput_settings = list(printT1=1))
# modelOutput_settings = list(printT1=1) adds t-values for the test whether the coefficient is significantly different from 1 - that's relevant for the lambdas! A lambda equal to 1 would mean that there actually is no nest.

apollo_lrTest(model, "MNL_Exam_question3")

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)
```
f) To test for different levels of correlations between the tastes in each nest, we need to estimate a mixed logit model, so the parameter $$ 0 \le \lambda \le 1$$ , which  measures the substitutability of products within a group, must be introduced.


